---
title: "Temporal Crime Trends at the Johns Hopkins East Baltimore Campus"
author: "Shannon Wongvibulsin"
date: "October 23, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load Necessary Libraries
Note: use ‘ggplot2’ 2.2.0 (add note to readme file)
```{r}
packages<-c("lubridate", "dplyr","ggplot2","reshape2", "xtable")

for (i in packages){
  if(!require(i,character.only = T,quietly=T,warn.conflicts = F)){
    install.packages(i)
  }
  require(i,character.only = T,quietly=T,warn.conflicts = F)
}

library(lubridate)
library(dplyr)
library(ggplot2)
library(reshape2)
library(xtable)
library(chron)
library(magrittr)

```

## Inport data and look at structure of data
Data from Johns Hopkins University Clery Compliance Administrator after conversion to CSV formats.

```{r data}
#setwd("/Users/purplesw/Dropbox/final_project/data")
#setwd("~/final_project/data")
setwd("./data")
dat.15 <- read.csv("2015crimelog.csv", stringsAsFactor = FALSE)
dat.16 <- read.csv("2016crimelog.csv", stringsAsFactor = FALSE)
dat.17 <- read.csv("2017crimelog.csv", stringsAsFactor = FALSE)

str(dat.15)
str(dat.16)
str(dat.17)

```

## Crime categories in different years
```{r crime categories}
unique(dat.15$Crime)
unique(dat.16$Crime)
unique(dat.17$Crime)
```

## Clean data 
```{r clean}

# remove blank entries
dat.15 <- dat.15[!dat.15$Crime == "",]
dat.16 <- dat.16[!dat.16$Crime == "",]
dat.17 <- dat.17[!dat.17$Crime == "",]

dat.16 <- dat.16[!dat.16$Crime == "21",]
dat.17 <- dat.17[!dat.17$Crime == "18",]
```


## Format data 
create time of day, day of week, month, season, year
```{r format}
# Data for 2015
# extract time information (time without AM, PM)
dat.15$time <- sapply(dat.15$Date.Time.Occurred, function(x) strsplit(as.character(x), " ")[[1]][2])
# AM, PM time information
dat.15$ap <- sapply(dat.15$Date.Time.Occurred, function(x) strsplit(as.character(x), " ")[[1]][3])
# combine time and AM, PM
dat.15$timeap <- paste(dat.15$time, dat.15$ap) 
# convert to military time
dat.15$mtime <- substr(strptime(dat.15$timeap, "%I:%M %p"), 11,19) 
# remove white space
dat.15$mtime <- trimws(dat.15$mtime)
# extract hour information
dat.15$hours <- as.numeric(substr(dat.15$mtime, 1,2))

# categorize times into time of day (tod)
dat.15$ctime <- hour(hms(dat.15$mtime)) 
tbreaks <- hour(hm("00:00", "5:00", "11:00", "17:00", "23:59"))
tlabels <- c("Night", "Morning", "Afternoon", "Evening")
dat.15$tod <- cut(x=dat.15$ctime, breaks = tbreaks, labels = tlabels, include.lowest = TRUE)

# date, day, month
dat.15$date <- sapply(dat.15$Date.Time.Reported, function(x) strsplit(as.character(x), " ")[[1]][1]) %>% mdy()
dat.15$day <- weekdays(dat.15$date)
dat.15$day <- factor(dat.15$day, levels = c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday", "Sunday"))
dat.15$month <- months(dat.15$date)
dat.15$month <- factor(dat.15$month, levels = month.name)

# table crimes by tod
crime.tod.15 <- as.data.frame(table(dat.15$Crime, dat.15$tod))

# Data for 2016
# extract time information (time without AM, PM)
dat.16$time <- sapply(dat.16$Date.Time.Occurred, function(x) strsplit(as.character(x), " ")[[1]][2]) 
# AM, PM time information
dat.16$ap <- sapply(dat.16$Date.Time.Occurred, function(x) strsplit(as.character(x), " ")[[1]][3]) 
# combine time and AM, PM
dat.16$timeap <- paste(dat.16$time, dat.16$ap) 
# convert to military time
dat.16$mtime <- substr(strptime(dat.16$timeap, "%I:%M %p"), 11,19) 
# remove white space
dat.16$mtime <- trimws(dat.16$mtime)
# extract hour information
dat.16$hours <- as.numeric(substr(dat.16$mtime, 1,2))

# categorize times into time of day (tod)
dat.16$ctime <- hour(hms(dat.16$mtime)) 
tbreaks <- hour(hm("00:00", "5:00", "11:00", "17:00", "23:59"))
tlabels <- c("Night", "Morning", "Afternoon", "Evening")
dat.16$tod <- cut(x=dat.16$ctime, breaks = tbreaks, labels = tlabels, include.lowest = TRUE)

# table crimes by tod
crime.tod.16 <- as.data.frame(table(dat.16$Crime, dat.16$tod))

# date, day, month
dat.16$date <- sapply(dat.16$Date.Time.Reported, function(x) strsplit(as.character(x), " ")[[1]][1]) %>% mdy()
dat.16$day <- weekdays(dat.16$date)
dat.16$day <- factor(dat.16$day, levels = c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday", "Sunday"))
dat.16$month <- months(dat.16$date)
dat.16$month <- factor(dat.16$month, levels = month.name)


# Data for 2017
# extract time information (time without AM, PM)
dat.17$time <- sapply(dat.17$Date.Time.Occurred, function(x) strsplit(as.character(x), " ")[[1]][2])
# AM, PM time information
dat.17$ap <- sapply(dat.17$Date.Time.Occurred, function(x) strsplit(as.character(x), " ")[[1]][3]) 
# combine time and AM, PM
dat.17$timeap <- paste(dat.17$time, dat.17$ap) # combine time and AM, PM
# convert to military time
dat.17$mtime <- substr(strptime(dat.17$timeap, "%I:%M %p"), 11,19) 
# remove white space
dat.17$mtime <- trimws(dat.17$mtime)
dat.17$hours <- as.numeric(substr(dat.17$mtime, 1,2))

# categorize times into time of day (tod)
dat.17$ctime <- hour(hms(dat.17$mtime)) 
tbreaks <- hour(hm("00:00", "5:00", "11:00", "17:00", "23:59"))
tlabels <- c("Night", "Morning", "Afternoon", "Evening")
dat.17$tod <- cut(x=dat.17$ctime, breaks = tbreaks, labels = tlabels, include.lowest = TRUE)

# table crimes by tod
crime.tod.17 <- as.data.frame(table(dat.17$Crime, dat.17$tod))

# date, day, month
dat.17$date <- sapply(dat.17$Date.Time.Reported, function(x) strsplit(as.character(x), " ")[[1]][1]) %>% mdy()
dat.17$day <- weekdays(dat.17$date)
dat.17$day <- factor(dat.17$day, levels = c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday", "Sunday"))

dat.17$month <- months(dat.17$date)
dat.17$month <- factor(dat.17$month, levels = month.name)


# create function for summarizing total crime vs. variable of interest
crime.sum <- function(df, x){
  dft <- as.data.frame(table(df$Crime, df[,x]))
  dft <- melt(dft, c("Var1", "Var2"))
  g.dft <- dft %>% group_by(Var2) %>% summarize(n = sum(value))
  dg.dft <- as.data.frame(g.dft)
  
  p <- ggplot(data = dg.dft, aes(x=Var2, y=n)) + theme(axis.text.x = element_text(angle = 90, hjust = 1)) + stat_smooth((aes(group = 1)), se = FALSE)
  return(p)
  
}

# function for summarizing total crime vs. month by year
crime.sum.yr <- function(df15, df16, g.dft.17, x){
  dft.15 <- as.data.frame(table(df15$Crime, df15[,x]))
  dft.16 <- as.data.frame(table(df16$Crime, df16[,x]))
  dft.15 <- melt(dft.15, c("Var1", "Var2"))
  dft.16 <- melt(dft.16, c("Var1", "Var2"))
  g.dft.15 <- dft.15 %>% group_by(Var2) %>% summarize(n = sum(value))
  g.dft.16 <- dft.16 %>% group_by(Var2) %>% summarize(n = sum(value))
  g.dft.15 <- g.dft.15 %>% mutate(year = 2015)
  g.dft.16 <- g.dft.16 %>% mutate(year = 2016)
  dg.dft.15 <- as.data.frame(g.dft.15)
  dg.dft.16 <- as.data.frame(g.dft.16)
  m.df <- rbind(g.dft.15, g.dft.16)
  m.df <- rbind(m.df, g.dft.17)
  
  p <- ggplot(data = m.df, aes(x=Var2, y=n, group = factor(year), color = factor(year))) + stat_smooth(se = FALSE) + theme(axis.text.x = element_text(angle = 90, hjust = 1), legend.position = c(0.8,0.9), legend.title = element_blank()) + ylab("Crime Count") + xlab(NULL) + ggtitle("Total Crime by Month")
  return(p)
}


# function for summarizing total crime vs. hour by year
crime.h <- function(df15, df16, df17, x){
  dft.15 <- as.data.frame(table(df15$Crime, df15[,x]))
  dft.16 <- as.data.frame(table(df16$Crime, df16[,x]))
  dft.17 <- as.data.frame(table(df17$Crime, df17[,x]))
  dft.15 <- melt(dft.15, c("Var1", "Var2"))
  dft.16 <- melt(dft.16, c("Var1", "Var2"))
  dft.17 <- melt(dft.17, c("Var1", "Var2"))
  g.dft.15 <- dft.15 %>% group_by(Var2) %>% summarize(n = sum(value))
  g.dft.16 <- dft.16 %>% group_by(Var2) %>% summarize(n = sum(value))
  g.dft.17 <- dft.17 %>% group_by(Var2) %>% summarize(n = sum(value))
  g.dft.15 <- g.dft.15 %>% mutate(year = 2015)
  g.dft.16 <- g.dft.16 %>% mutate(year = 2016)
  g.dft.17 <- g.dft.17 %>% mutate(year = 2017)
  dg.dft.15 <- as.data.frame(g.dft.15)
  dg.dft.16 <- as.data.frame(g.dft.16)
  dg.dft.17 <- as.data.frame(g.dft.17)
  m.df <- rbind(g.dft.15, g.dft.16)
  m.df <- rbind(m.df, g.dft.17)
  
  p <- ggplot(data = m.df, aes(x=Var2, y=n, group = factor(year), color = factor(year))) + stat_smooth(se = FALSE) + theme(axis.text.x = element_text(angle = 90, hjust = 1), legend.position = c(0.8,0.9), legend.title = element_blank()) + ylab("Crime Count") + xlab("Hours in Military Time") + ggtitle("Total Crime by Hour")
  return(p)
  
}

# function for summarizing total crime vs. time of day by year
crime.t <- function(df15, df16, df17, x){
  dft.15 <- as.data.frame(table(df15$Crime, df15[,x]))
  dft.16 <- as.data.frame(table(df16$Crime, df16[,x]))
  dft.17 <- as.data.frame(table(df17$Crime, df17[,x]))
  dft.15 <- melt(dft.15, c("Var1", "Var2"))
  dft.16 <- melt(dft.16, c("Var1", "Var2"))
  dft.17 <- melt(dft.17, c("Var1", "Var2"))
  g.dft.15 <- dft.15 %>% group_by(Var2) %>% summarize(n = sum(value))
  g.dft.16 <- dft.16 %>% group_by(Var2) %>% summarize(n = sum(value))
  g.dft.17 <- dft.17 %>% group_by(Var2) %>% summarize(n = sum(value))
  g.dft.15 <- g.dft.15 %>% mutate(year = 2015)
  g.dft.16 <- g.dft.16 %>% mutate(year = 2016)
  g.dft.17 <- g.dft.17 %>% mutate(year = 2017)
  dg.dft.15 <- as.data.frame(g.dft.15)
  dg.dft.16 <- as.data.frame(g.dft.16)
  dg.dft.17 <- as.data.frame(g.dft.17)
  m.df <- rbind(g.dft.15, g.dft.16)
  m.df <- rbind(m.df, g.dft.17)
  
  p <- ggplot(data = m.df, aes(x=Var2, y=n, group = factor(year), color = factor(year))) + stat_smooth(se = FALSE) + theme(axis.text.x = element_text(angle = 90, hjust = 1), legend.position = c(0.8,0.9), legend.title = element_blank()) + ylab("Crime Count") + xlab(NULL) + ggtitle("Total Crime by Time of Day")
  return(p)
  
}

# function for summarizing total crime vs. day of week by year
crime.d <- function(df15, df16, df17, x){
  dft.15 <- as.data.frame(table(df15$Crime, df15[,x]))
  dft.16 <- as.data.frame(table(df16$Crime, df16[,x]))
  dft.17 <- as.data.frame(table(df17$Crime, df17[,x]))
  dft.15 <- melt(dft.15, c("Var1", "Var2"))
  dft.16 <- melt(dft.16, c("Var1", "Var2"))
  dft.17 <- melt(dft.17, c("Var1", "Var2"))
  g.dft.15 <- dft.15 %>% group_by(Var2) %>% summarize(n = sum(value))
  g.dft.16 <- dft.16 %>% group_by(Var2) %>% summarize(n = sum(value))
  g.dft.17 <- dft.17 %>% group_by(Var2) %>% summarize(n = sum(value))
  g.dft.15 <- g.dft.15 %>% mutate(year = 2015)
  g.dft.16 <- g.dft.16 %>% mutate(year = 2016)
  g.dft.17 <- g.dft.17 %>% mutate(year = 2017)
  dg.dft.15 <- as.data.frame(g.dft.15)
  dg.dft.16 <- as.data.frame(g.dft.16)
  dg.dft.17 <- as.data.frame(g.dft.17)
  m.df <- rbind(g.dft.15, g.dft.16)
  m.df <- rbind(m.df, g.dft.17)
  
  p <- ggplot(data = m.df, aes(x=Var2, y=n, group = factor(year), color = factor(year))) + stat_smooth(se = FALSE) + theme(axis.text.x = element_text(angle = 90, hjust = 1), legend.position = c(0.8,0.9), legend.title = element_blank()) + ylab("Crime Count") + xlab(NULL) + ggtitle("Total Crime by Day")
  return(p)
  
}

  dft.15 <- as.data.frame(table(dat.15$Crime, dat.15[,"month"]))
  dft.16 <- as.data.frame(table(dat.16$Crime, dat.16[,"month"]))
  dft.15 <- melt(dft.15, c("Var1", "Var2"))
  dft.16 <- melt(dft.16, c("Var1", "Var2"))
  g.dft.15 <- dft.15 %>% group_by(Var2) %>% summarize(n = sum(value))
  g.dft.16 <- dft.16 %>% group_by(Var2) %>% summarize(n = sum(value))
  g.dft.15 <- g.dft.15 %>% mutate(year = 2015)
  g.dft.16 <- g.dft.16 %>% mutate(year = 2016)
  dg.dft.15 <- as.data.frame(g.dft.15)
  dg.dft.16 <- as.data.frame(g.dft.16)
  m.df <- rbind(dg.dft.15, dg.dft.16)
  #m.df <- left_join(g.dft.15, g.dft.16, by = "Var2")

dcrime <- crime.d(dat.15, dat.16, dat.17, "day")
tcrime <- crime.t(dat.15, dat.16, dat.17, "tod")
hcrime <- crime.h(dat.15, dat.16, dat.17, "hours")
mcrime <- crime.sum.yr(dat.15, dat.16, dg.dft.17, "month")

plot_grid(mcrime, dcrime, hcrime, tcrime, labels = c("A", "B", "C", "D"))
```

## Calendar heatmap
```{r heatmap}

# tally the total number of crimes on each day
dat.15.total <- dat.15 %>% group_by(date) %>% tally()
dat.16.total <- dat.16 %>% group_by(date) %>% tally()
dat.17.total <- dat.17 %>% group_by(date) %>% tally()
total.yr <- rbind(dat.15.total, dat.16.total, dat.17.total)

require(RColorBrewer)
red_color_ramp = brewer.pal(9, "Reds")
ct <- calendarHeat(total.yr$date, total.yr$n,
varname="Daily Crime", color="red_color_ramp")

```


## Statistical testing

```{r Tukey}

# Tukey test to determine what the effect of time of day on the total number of crime during that time
total.15 <- group_by(dat.15, tod) %>%
  summarize(
    count = n()
  )

total.16 <- group_by(dat.16, tod) %>%
  summarize(
    count = n()
  )

total.17 <- group_by(dat.17, tod) %>%
  summarize(
    count = n()
  )


total.tod <- rbind(total.15, total.16, total.17)
total.aov <- aov(count ~ tod, data = total.tod)
TH.tod <- TukeyHSD(total.aov)


# Tukey test to determine what the effect of day of the week on the total number of crime on that day

day.15 <- group_by(dat.15, day) %>%
  summarize(
    count = n()
  )

day.16 <- group_by(dat.16, day) %>%
  summarize(
    count = n()
  )

day.17 <- group_by(dat.17, day) %>%
  summarize(
    count = n()
  )


total.day <- rbind(day.15, day.16, day.17)
day.aov <- aov(count ~ day, data = total.day)
TH.day <- TukeyHSD(day.aov)

print(xtable(TH.tod$tod))
print(xtable(TH.day$day))


# function to plot effect size and 95% CI
plotTukeyHSD <- plotTukeysHSD <- function(tukey.out,
                           x.axis.label = "Comparison",
                           y.axis.label = "Effect Size",
                       axis.adjust = 0,
                       adjust.x.spacing = 5){
  
  tukey.out <- as.data.frame(tukey.out[[1]])
  means <- tukey.out$diff
  categories <- row.names(tukey.out)
  groups <- length(categories)
  ci.low <- tukey.out$lwr
  ci.up  <- tukey.out$upr                         
  
  n.means <- length(means)
   
  #determine where to plot points along x-axis
  x.values <- 1:n.means
  x.values <- x.values/adjust.x.spacing
  
                             
  # calculate values for plotting limits            
  y.max <- max(ci.up) +                    
    max(ci.up)*axis.adjust
  y.min <- min(ci.low) - 
    max(ci.low)*axis.adjust
  
  if(groups == 2){ x.values <- c(0.25, 0.5)}
  if(groups == 3){ x.values <- c(0.25, 0.5,0.75)}
  
  x.axis.min <- min(x.values)-0.05
  x.axis.max <- max(x.values)+0.05
  
  x.limits <- c(x.axis.min,x.axis.max)
  
  #Plot means
  plot(means ~ x.values,
       xlim = x.limits,
       ylim = c(y.min,y.max),
       xaxt = "n",
       xlab = "",
       ylab = "",
       cex = 1.25,
       pch = 16)
  
  axis(side = 1, 
       at = x.values,
       labels = categories
      )
  
  #Plot upper error bar 
  lwd. <- 2
  arrows(y0 = means,
         x0 = x.values,
         y1 = ci.up,
         x1 = x.values,
         length = 0,
         lwd = lwd.)
  
  #Plot lower error bar
  arrows(y0 = means,
         x0 = x.values,
         y1 = ci.low,
         x1 = x.values,
         length = 0,
         lwd = lwd.) 
  
  #add reference line at 0
  abline(h = 0, col = 2, lwd = 2, lty =2)
  
  #mtext(text = x.axis.label,side = 1,line = 1.75)
  mtext(text = y.axis.label,side = 2,line = 1.95)
  # Error bars = 95% CI
  # mtext(text = "Effect Size of Time of Day on Total Crime",side = 3,line = 1)
  mtext(text = "Effect Size of Day of Week on Total Crime",side = 3,line = 1)
  
}      


plotTukeyHSD(TH.tod)
plotTukeyHSD(TH.day)

```


## create tables
```{r}
library(xtable)
options(xtable.floating = FALSE)

c.types.15 <- table(dat.15$Crime)

x.c.types.15 <- xtable(c.types.15)
print(xtable(c.types.15, caption = "Crime Types and Frequency"), caption.placement = "top")

c.types.16 <- table(dat.16$Crime)
print(xtable(c.types.16))

c.types.17 <- table(dat.17$Crime)
print(xtable(c.types.17))

```






