---
title: "crime_analysis"
author: "Shannon Wongvibulsin"
date: "September 11, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Johns Hopkins Security Alert - Crime Analysis & Visualization

## Inport Data
Data from Johns Hopkins University Clery Compliance Administrator after conversion to CSV formats.

```{r data}
#setwd("/Users/purplesw/Dropbox/final_project/data")
#setwd("~/final_project/data")
setwd("./data")
dat.15 <- read.csv("2015crimelog.csv")
dat.16 <- read.csv("2016crimelog.csv")
dat.17 <- read.csv("2017crimelog.csv")
```

## Differences in crime types in different years
```{r}
unique(dat.15$Crime)
unique(dat.16$Crime)
unique(dat.17$Crime)
```


## Create plots
```{r}

t15 <- table(dat.15$Crime)
d15 <- as.data.frame(t15)
d15$Year <- "2015"
d15 <- d15[-1,]

dat.16 <- dat.16[!dat.16$Crime=="",]
dat.16 <- dat.16[!dat.16$Crime=="21",]
t16 <- table(dat.16$Crime)
d16 <- as.data.frame(t16)
d16$Year <- "2016"
d16 <- d16[-1:-2,]

dat.17 <- dat.17[!dat.17$Crime=="",]
dat.17 <- dat.17[!dat.17$Crime=="18",]
t17 <- table(dat.17$Crime)

d17 <- as.data.frame(t17)
d17$Year <- "2017"
d17 <- d17[-1:-2,]

total <- rbind(d15, d16)
total <- rbind(total, d17)


library(ggplot2)
#library(plotly)

p <- ggplot(data = total, aes(x=Var1, y=Freq, fill = Year)) + geom_bar(stat="identity", position=position_dodge()) + theme(axis.text.x = element_text(angle = 90, hjust = 1), plot.title = element_text(hjust = 0.5, face = "bold"), legend.position = c(0.95, 0.85), axis.title = element_text(face = "bold")) + labs(x = "Crime Type", y = "Number of Incidents") + ggtitle("Crime Frequency by Year")

pp <- ggplotly(p) 

pp


```

## Dates
consider season, month, day of week, time

```{r}
library(lubridate)
library(dplyr)
library(reshape2)

#dat.15$day <- weekdays(as.Date(dat.15$Date.Time.Reported))
#dat.15$month <- months(as.Date(dat.15$Date.Time.Reported))
#dat.15$quarter <- quarters(as.Date(dat.15$Date.Time.Reported))

# extract time information from Date.Time.Occurred; if strsplit(as.character(x), " ")[[1]][3]  == "PM", add 12 hours except  

#dat.15$dt <- sapply(dat.15$Date.Time.Occurred, function(x) as.POSIXlt(as.character(x), format = "%m/%d/%Y %I:%M %p", tz = "UTC"))


dat.15$timehour <- sapply(dat.15$Date.Time.Occurred, function(x) strsplit(as.character(x), " ")[[1]][2]) 
dat.15$timeap <- sapply(dat.15$Date.Time.Occurred, function(x) strsplit(as.character(x), " ")[[1]][3])

dat.15$times <- paste(dat.15$timehour, dat.15$timeap)
dat.15$mtime <- substr(strptime(dat.15$times, "%I:%M %p"), 11,19)
dat.15$mtime <- trimws(dat.15$mtime)

time <- hour(hms(dat.15$mtime))
tbreaks <- hour(hm("00:00", "6:00", "12:00", "18:00", "23:59"))
tlabels <- c("Night", "Morning", "Afternoon", "Evening")
dat.15$tod <- cut(x=time, breaks = tbreaks, labels = tlabels, include.lowest = TRUE)

crime.tod.15 <- as.data.frame(table(dat.15$Crime, dat.15$tod))
crime.tod.15 <- crime.tod.15[-c(1, 15, 29, 43),]

pd <- ggplot(data = crime.tod.15, aes(x=Var1, y=Freq, fill = Var2)) + geom_bar(stat="identity", position=position_dodge()) + theme(axis.text.x = element_text(angle = 90, hjust = 1), plot.title = element_text(hjust = 0.5, face = "bold"), legend.position = c(0.95, 0.85), axis.title = element_text(face = "bold")) + labs(x = "Crime Type", y = "Number of Incidents") + ggtitle("Crime Frequency by Time of Day for 2015") + guides(fill = guide_legend(title = "Time of Day"))

dat.16$timehour <- sapply(dat.16$Date.Time.Occurred, function(x) strsplit(as.character(x), " ")[[1]][2]) 
dat.16$timeap <- sapply(dat.16$Date.Time.Occurred, function(x) strsplit(as.character(x), " ")[[1]][3])

dat.16$times <- paste(dat.16$timehour, dat.16$timeap)
dat.16$mtime <- substr(strptime(dat.16$times, "%I:%M %p"), 11,19)
dat.16$mtime <- trimws(dat.16$mtime)

time <- hour(hms(dat.16$mtime))
tbreaks <- hour(hm("00:00", "6:00", "12:00", "18:00", "23:59"))
tlabels <- c("Night", "Morning", "Afternoon", "Evening")
dat.16$tod <- cut(x=time, breaks = tbreaks, labels = tlabels, include.lowest = TRUE)

crime.tod.16 <- as.data.frame(table(dat.16$Crime, dat.16$tod))

dat.17$timehour <- sapply(dat.17$Date.Time.Occurred, function(x) strsplit(as.character(x), " ")[[1]][2]) 
dat.17$timeap <- sapply(dat.17$Date.Time.Occurred, function(x) strsplit(as.character(x), " ")[[1]][3])

dat.17$times <- paste(dat.17$timehour, dat.17$timeap)
dat.17$mtime <- substr(strptime(dat.17$times, "%I:%M %p"), 11,19)
dat.17$mtime <- trimws(dat.17$mtime)

time <- hour(hms(dat.17$mtime))
tbreaks <- hour(hm("00:00", "6:00", "12:00", "18:00", "23:59"))
tlabels <- c("Night", "Morning", "Afternoon", "Evening")
dat.17$tod <- cut(x=time, breaks = tbreaks, labels = tlabels, include.lowest = TRUE)

crime.tod.17 <- as.data.frame(table(dat.16$Crime, dat.16$tod))







#dat.15$timeday[dat.15$timehour]

dat.15$date <- sapply(dat.15$Date.Time.Reported, function(x) strsplit(as.character(x), " ")[[1]][1]) %>% mdy()
dat.15$day <- weekdays(dat.15$date)
dat.15$month <- months(dat.15$date)
dat.15$quarter <- quarters(dat.15$date)

dat.15 <- dat.15[-288,]

dw15 <- as.data.frame(table(dat.15$Crime, dat.15$day))

dwm15 <- melt(dw15, c("Var1", "Var2"))

dwm15.s <- dwm15[dwm15$Var1== "Assault" | dwm15$Var1== "Theft", ]

pd <- ggplot(data = dwm15, aes(x=Var1, y=value, fill = Var2)) + geom_bar(stat="identity", position=position_dodge()) + theme(axis.text.x = element_text(angle = 90, hjust = 1), legend.title=element_blank(), plot.title = element_text(hjust = 0.5, face = "bold"), axis.title = element_text(face = "bold")) + scale_fill_discrete(breaks = c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday", "Sunday")) + ggtitle("Crime Frequency by Day of Week for 2015") + labs(x = "Crime Type", y = "Number of Incidents")

pd.s <- ggplot(data = dwm15.s, aes(x=Var1, y=value, fill = Var2)) + geom_bar(stat="identity", position=position_dodge()) + theme(axis.text.x = element_text(angle = 90, hjust = 1), legend.title=element_blank(), plot.title = element_text(hjust = 0.5, face = "bold"), axis.title = element_text(face = "bold")) + scale_fill_discrete(breaks = c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday", "Sunday")) + ggtitle("Crime Frequency by Day of Week for 2015") + labs(x = "Crime Type", y = "Number of Incidents")

ppd <- ggplotly(pd) 

ppd

dw15m <- as.data.frame(table(dat.15$Crime, dat.15$month))

dwm15m <- melt(dw15m, c("Var1", "Var2"))

dwm15m.s <- dwm15m[dwm15m$Var1== "Assault" | dwm15m$Var1== "Theft", ]

#pm <- ggplot(data = dwm15m, aes(x=Var1, y=value, fill = Var2)) + geom_bar(stat="identity", position=position_dodge()) + theme(axis.text.x = element_text(angle = 90, hjust = 1))

pm <- ggplot(data = dwm15m, aes(x=Var1, y=value, fill = Var2)) + geom_bar(stat="identity", position=position_dodge()) + theme(axis.text.x = element_text(angle = 90, hjust = 1), legend.title=element_blank(), plot.title = element_text(hjust = 0.5, face = "bold"), axis.title = element_text(face = "bold")) + scale_fill_discrete(breaks = c("January", "February", "March", "April", "May", "June", "July", "August", "September", "October", "November", "December")) + ggtitle("Crime Frequency by Month for 2015") + labs(x = "Crime Type", y = "Number of Incidents")

pm.s <- ggplot(data = dwm15m.s, aes(x=Var1, y=value, fill = Var2)) + geom_bar(stat="identity", position=position_dodge()) + theme(axis.text.x = element_text(angle = 90, hjust = 1), legend.title=element_blank(), plot.title = element_text(hjust = 0.5, face = "bold"), axis.title = element_text(face = "bold")) + scale_fill_discrete(breaks = c("January", "February", "March", "April", "May", "June", "July", "August", "September", "October", "November", "December")) + ggtitle("Crime Frequency by Month for 2015") + labs(x = "Crime Type", y = "Number of Incidents")

ppm <- ggplotly(pm) 

ppm

dw15q <- as.data.frame(table(dat.15$Crime, dat.15$quarter))

dwm15q <- melt(dw15q, c("Var1", "Var2"))

pq <- ggplot(data = dwm15q, aes(x=Var1, y=value, fill = Var2)) + geom_bar(stat="identity", position=position_dodge()) + theme(axis.text.x = element_text(angle = 90, hjust = 1))

ppq <- ggplotly(pq) 

ppq

# seasons

newdat.15$num.month <- match(newdat.15$month, month.name)

newdat.15 <-
newdat.15 %>%
  mutate(Season = ifelse(num.month >= 3 & num.month <= 5, "Spring", 
                          ifelse(num.month >= 6 & num.month <= 8, "Summer",
                            ifelse(num.month >= 9 & num.month <= 11, "Autumn", "Winter"))))

dw15s <- as.data.frame(table(newdat.15$Crime, newdat.15$Season))

dwm15s <- melt(dw15s, c("Var1", "Var2"))

#ps <- ggplot(data = dwm15s, aes(x=Var1, y=value, fill = Var2)) + geom_bar(stat="identity", position=position_dodge()) + theme(axis.text.x = element_text(angle = 90, hjust = 1))

ps <- ggplot(data = dwm15s, aes(x=Var1, y=value, fill = Var2)) + geom_bar(stat="identity", position=position_dodge()) + theme(axis.text.x = element_text(angle = 90, hjust = 1), legend.title=element_blank(), plot.title = element_text(hjust = 0.5, face = "bold"), axis.title = element_text(face = "bold")) + scale_fill_discrete(breaks = c("Spring", "Summer", "Autumn", "Winter")) + ggtitle("Crime Frequency by Season for 2015") + labs(x = "Crime Type", y = "Number of Incidents")

pps <- ggplotly(ps) 

pps

```

## Statistical Testing: ANOVA
```{r}

dat.15.theft <- filter(dat.15, dat.15$Crime == "Theft")
dat.16.theft <- filter(dat.16, dat.16$Crime == "Theft")
dat.17.theft <- filter(dat.17, dat.17$Crime == "Theft")

test.15 <- group_by(dat.15.theft, tod) %>%
  summarize(
    count = n()
  )

test.16 <- group_by(dat.16.theft, tod) %>%
  summarize(
    count = n()
  )

test.17 <- group_by(dat.17.theft, tod) %>%
  summarize(
    count = n()
  )

# NA values
test.16 <- test.16[1:4,]
test.17 <- test.17[1:4,]

test <- rbind(test.15, test.16, test.17)

test.aov <- aov(count ~ tod, data = test)
TH <- TukeyHSD(test.aov)
pt <- pairwise.t.test(test$count, test$tod)

print(xtable(TH$tod))

print(xtable(pt$p.value))

```



## Location
Need to do text processing of general locations (buildings, areas)
155 unique locations 
look at groupings of general areas
look at correlations of crime types, time and areas
```{r}
t.15 <- table(dat.15$General.Location)
d.15 <- as.data.frame(t.15)

library("dataframes2xls")
write.xls(d.15, "locationlist.xls")

write.xls(d.15, "locationtest.xls", quote = FALSE)

library("stringr")
d.15$outside <- str_detect(d.15$Var1, "block")

dat.15$outside <- str_detect(dat.15$General.Location, "block")

```

## Geocode Locations
```{r}
jh.map <- qmap("johns hopkins hospital", zoom = 14)

jh.map + geom_point(aes(x=lon, y=lat, colour = name), data = h.df) + ggtitle("Map of Crime at the Johns Hopkins East Baltimore Medical Campus")


h.df <- geocode(c("1550 Orleans St, Baltimore", "1800 Orleans St, Baltimore"))
h.df$name <- c("CRB", "Main Hospital")
g.loc <- paste(h.df$lat, h.df$lon, sep = ":")
m.google <- data.frame(h.df, g.loc)
g.inter.map <- gvisMap(data = m.google, locationvar = "g.loc", tipvar = "name")
plot(g.inter.map)


# cite: http://www.storybench.org/geocode-csv-addresses-r/
setwd("/Users/purplesw/Dropbox/final_project/data")
origAddress <- read.csv("location.csv", stringsAsFactors = FALSE)

a.df <- as.data.frame(table(origAddress$addresses))

for(i in 1:nrow(a.df)){
  temp <- geocode(as.character(a.df$Var1[i]))
  a.df$lon[i] <- temp$lon
  a.df$lat[i] <- temp$lat
  Sys.sleep(2)
}


geocoded <- data.frame(stringsAsFactors = FALSE)

# Loop through the addresses to get the latitude and longitude of each address and add it to the
# origAddress data frame in new columns lat and lon
for(i in 1:nrow(origAddress))
{
  # Print("Working...")
  result <- geocode(origAddress$addresses[i], output = "latlona", source = "google")
  origAddress$lon[i] <- as.numeric(result[1])
  origAddress$lat[i] <- as.numeric(result[2])
  origAddress$geoAddress[i] <- as.character(result[3])
  Sys.sleep(2)
}
# Write a CSV file containing origAddress to the working directory
write.csv(origAddress, "geocoded.csv", row.names=FALSE)
write.csv(a.df, "uniquegeocode.csv", row.names=FALSE)

#setwd("/Users/purplesw/Dropbox/final_project/code")
g.df <- read.csv("uniquegeocode.csv", stringsAsFactors = FALSE)

colnames(g.df)[1] <- "addresses"
geo.loc <- left_join(origAddress, g.df, by  = "addresses")

dat.15.a <- dat.15
colnames(dat.15.a)[5] <- "Var1"

for(i in 1:nrow(geo.loc)){
  geo.loc$Var1[i] <- noquote(geo.loc$Var1[i])
}


geo.loc.t <- left_join(dat.15.a, geo.loc, by = "Var1")



geo.loc.t.df <- as.data.frame(table(geo.loc.t$addresses))

colnames(geo.loc.t.df)[1] <- "addresses"

g.df.t <- left_join(geo.loc.t.df, g.df, by = "addresses")

plot(g.df$lon, g.df$lat, xlab = "longitude", ylab = "latitude", main = "Longitude and Latitude of Crime Locations")

jhu.map <- qmap("johns hopkins hospital", zoom = 16)
jhu.map + geom_point(aes(x = lon, y = lat, colour = factor(Freq)), size = 2, data = g.df)

jhu.1 + geom_point(aes(x = lon, y = lat, colour = factor(Freq.x)), size = 2, data = g.df.t)


jhu.1 + geom_point(aes(x = lon, y = lat, colour = factor(Freq)), size = 3, data = g.df) + scale_fill_discrete(name = "Crime Incidents \nat Each Location")

#plot(g.df$lon, g.df$lat, xlab = "lon", ylab = "lat", xlim = c(-76.59734, -76.5), ylim = c(39.2, 39.3))
```

## create tables
```{r}
library(xtable)
options(xtable.floating = FALSE)

c.types.15 <- table(dat.15$Crime)

x.c.types.15 <- xtable(c.types.15)
print(xtable(c.types.15, caption = "Crime Types and Frequency"), caption.placement = "top")

c.types.16 <- table(dat.16$Crime)
print(xtable(c.types.16))

c.types.17 <- table(dat.17$Crime)
print(xtable(c.types.17))


```



## To Do
- parse time of day from "Date.Time.Occurred"
- rearrange labels for plots
- add title and axis lables to plots

Analysis 
- crimes by location
- crimes by inside/outside
- crimes inside/outside by time of day
- crimes inside/outside by season





